---
title: "stl_lead_inequity"
author: "Julia Parish"
date: "8/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(janitor)
library(skimr)
library(here)

```

## Read in stl_blood_lead.csv data as stl_lead 

```{r}
stl_lead <- read_delim(here::here("data", "stl_blood_lead.csv"), delim = ";") %>% clean_names() %>% 
mutate(prop_white = white / total_pop * 100) %>% 
  mutate(pct_elevated = str_replace(pct_elevated, 
                                    pattern = ",",
                           replacement = ".")) %>% 
  mutate(pct_elevated = as.numeric(pct_elevated))

```

#Do some basic exploration of the dataset, and janitor::clean_names() if useful

```{r}
skimr::skim(stl_lead)

names(stl_lead)

```

```{r}
stl_lead <- stl_lead %>% clean_names()

```

##In a new code chunk, from stl_lead create a new data frame called stl_lead_prop that has one additional column called prop_white that returns the percent of each census tract identifying as white (variable white in the dataset divided by variable totalPop, times 100). You may need to do some Googling. Hint: dplyr::mutate(new_col = col_a / col_b) will create a new column new_col that contains the value of col_a / col_b

 see above for code but this is the answer:
 mutate(prop_white = white / total_pop * 100) %>% 
  mutate(pct_elevated = str_replace(pct_elevated, 
                                    pattern = ",",
                           replacement = ".")) %>% 
  mutate(pct_elevated = as.numeric(pct_elevated))

## In a new code chunk, create a scatterplot graph of the percentage of children in each census tract with elevated blood lead levels (pctElevated) versus the percent of each census tract identifying as white.
Customize by updating several aesthetics (e.g. size, opacity (see alpha =), color, etc.)
Store the scatterplot as stl_lead_plot
Have the scatterplot returned in the knitted html - customize the size that it appears when knitted
Also save a .png of the scatterplot to figs, with dimensions of (6 x 5) (width x height)
In text above or below the scatterplot, write 1 - 2 sentences describing the overall trend that you observe from your graph

```{r}
stl_lead_plot <- ggplot(data = stl_lead, aes(x = prop_white, y = pct_elevated)) + 
  geom_point(color = "violet", alpha= 0.5) + 
  labs(title = "Children with elevated blood levels",
       x = "% Population White",
       y = "Elevated PCT") 
  
  
stl_lead_plot
```
```{r}
ggsave(here::here("figs", "stl_lead_plot.png"), height = 6, width = 5)

stl_lead_plot
```

### Figure 1: As the population identifying as white increases the number of children with elevated PCT blood levels decreases

side note - 
# To reverse y axis - scale_y_reverse()
# To reverse x & y - coord_flip()

Create a histogram of only the pctElevated column in the data frame
Customize the fill, color, and size aesthetics - test some stuff! Feel free to make it awful.
Once you’ve played around with customization, export the histogram as a .jpg to the figs folder
Make sure the histogram also shows up in your knitted html (customize the size)

```{r}
stl_lead_histogram <- ggplot(stl_lead,
                      aes(x = pct_elevated)) + 
  geom_histogram(fill = "salmon") + 
  labs(title = "Elevated PCT Levels in Children",
       x = "PCT Levels",
       y = "Number of Children")

stl_lead_histogram

```

```{r}
ggsave(here::here("figs", "stl_lead_histogram.png"))


```
       
       
